An Innovative Storage Stack Addressing Extreme Scale Platforms and Big Data Applications

### 2014-04-30 paper.tex
minor fix to correctly reference IOD VOL image
added some to the related work and fixed up the intro some more. added some global images

The complete design seeks to offer high availability, byte-granular,
multi-version concurrency control. Through the use of a copy-on-write style
mechanism reducing space and time requirements, multiple versions of an object
can be stored efficiently. By assuming the client interface will be through an
IO library, a more complicated interface that offers richer functionality can
be incorporated while requiring only minimal end-user code changes.  Managing
most data access in a platform-local layer rather than requiring writing to
centralized storage will better support the performance and energy requirements
of extreme scale integrated application workflows. At a more detailed view, the
various layers of the IO stack each contribute different functionality and
performance implications.

@@

The architecture incorporates five layers, some of which are potentially
optional. The top layer is a high level IO library, such as the demonstration
HDF-5 library~\cite{hdf5}. The intent is to only have access to the storage
stack through such an API to manage the complexity of working with the lower
layers and to enable advanced functionality that would require more direct user
intervention. This layer incorporate the necessary features for ACGs from a
end-user's perspective.

Below the user API is an IO forwarding layer that redirects IO calls from the
compute nodes to the IO dispatching layer.  This IO forwarding layer is
analogous to the function of the IO nodes in a BlueGene machine or the passive
data staging processes demonstrated
previously~\cite{nisar:2008:staging,Abbasi:2009:datatap}. The next two layers
have considerable functionality. The IO dispatcher (IOD) serves as the primary
storage interface for the IO stack and is the only way to reach the persistent
storage array in lower layers. Ideally, the IOD layer's functionality can be
optional based on available hardware. Much of the functionality offered at this
layer would shift either up or down the stack as discussed in detail below. The
Distributed Asynchronous Object Storage (DAOS) layer serves as the persistent
storage interface and is intended to be the traditional file system-like
foundation on which everything else is built with no dependence on any
technologies specified above it. For example, the IOD layer with or without
burst buffers is not required for DAOS to operate properly.  Instead, the DAOS
layer can handle all of the IO operations from the user API layer, albiet with
the potential performance penalty of manipulating the shared, persistent
storage array. At the bottom is the Versioning Object Storage Device (VOSD).
It serves as the interface for storing objects of all types efficiently. Think
of this layer as the physical disk interface layer. In terms of Lustre, this
would replace the file system on individual storage targets with an interface
friendlier to the containers of objects concept used in the higher layers.

Along with the analysis of the published design documents, a discussion of the
design philosophy representing the overall intent is presented. This
information represents information that may or may not have been written down,
but is the intent of ultimate product. These insights were gained based on
personal conversations with the team
members~\cite{JohnBent,QuinceyKoziol,EricBarton} understanding the current
design as written. These ideas are presented to
give a fuller picture of where the project is going rather than dwelling on the
limitations of the published documents.

@@

The major previous work is really limited to full file systems rather than the
mountain of file system adjustments made over the years. The major players
related to this work include Ceph, PVFS, Lustre, GPFS, and Panasas.

EXPLAIN A BIT ABOUT HOW EACH CANNOT ADDRESS THE PROBLEMS WITH A CITATION FOR
EACH. DID I MISS A MAJOR FILE SYSTEM OF NOTE?

@@

\item
  \textbf{File}: In the HDF5 data model the container of an HDF5 infoset
  is represented by a file. It is a collection of objects that also
  explains the relationship between them. Every file begins with a root
  group ''/'', which serves as the ''starting-point'' in the object
  hierarchy.

@@

\section{IOD Layer}
\label{sec:iod}

@@

%Current NAND-based flash devices top out at around 400 MB/sec.
%The key spec that is missing from this number is that 400 MB/sec is a measure
%of the fixed number of available IOPS multiplied by the block size. This
%represents the ideal streaming performance possible. The problem is that it
%costs an IOP to read 1 byte or 1 block (4KB or 8KB, depending on the device).
%It costs 1 IOP to write a full block--usually. In some cases, it will cost 2
%IOPs.  This accounts for the required pre-erase write prior to writing to a
%reused block. In the worst case, it can be 3 IOPS per write. This would be for
%a partial block write (read the old block, erase, write the modified block).
%One-third of 400 MB/sec, about 133 MB/sec, is well below the streaming
%performance of HDDs.  Granted, there is still rotational and seek latency to
%deal with for HDDs, but the advantage for SSDs has evaporated and potentially
%turned into a penalty at a considerable cost premium.  There are faster SSD
%solutions on the market that incorporate DRAM for caching and using the PCIe
%bus, for example, but their price precludes them from use in an extreme scale
%platform.

@@

The second iteration of the protocol~\cite{lofstead:2013:pdsw-txn} fixed
scalability issues and demonstrated a scalable client-side coordination model
with excellent performance. The performance measured for a complex transaction
with \DDT is illustrated in Figure~\ref{fig:performance}. This performance is
explored in detail in a previous paper~\cite{lofstead:2013:pdsw-txn}.
%The
%breakdown of the number of participants in each role is shown in
%Table~\ref{tab:scaling}. For comparison, consider the Number of
%Sub-Coordinators equivalent to IOD processes. The Processes Per Sub-Coordinator
%represents the number of clients that use a particular IOD process. For these
%tests, we maintained a balanced distribution and always used at least two
%sub-coordinators to slow down the processing.

@@

%\begin{table}[ht]
%    \vspace{-0.15in}
%    \centering
%    \caption[Scaling Configuration]{Performance Tests Scaling Configuration}
%    \bigskip
%    \vspace{-0.15in}
%
%    \begin{tabular}{|r|r|r|}
%\hline
%Processes & \vtop{\hbox{\strut Number of}\hbox{\strut Sub-Coordinators}} & \vtop{\hbox{\strut Processes Per} \hbox{\strut Sub-Coordinator}}\\
%\hline
%%8 & 2 & 4 \\
%%16 & 2 & 8 \\
%%32 & 2 & 16 \\
%%64 & 2 & 32 \\
%%128 & 2 & 64 \\
%256 & 2 & 128 \\
%512 & 2 & 256 \\
%1024 & 4 & 256 \\
%2048 & 8 & 256 \\
%4096 & 16 & 256 \\
%8192 & 32 & 256 \\
%16384 & 64 & 256 \\
%32768 & 128 & 256 \\
%65536 & 256 & 256 \\
%\hline
%    \end{tabular}
%    \label{tab:scaling}
%\end{table}

@@

%\subsection{Comparison to Other Protocols}
%Alternatives, such as Paxos~\cite{Lamport:1998:paxos} algorithms like
%ZooKeeper~\cite{Hunt:2010:zookeeper}, suffer from two limitations making them
%unsuitable for this environment. First, the distributed servers in Paxos
%systems are all distributed copies of each other that eventually become
%consistent. Given the scale we wish to address, a single node's memory is
%unlikely to be able to hold all of the data necessary for many operations at
%scale. They also do not have a guarantee for when consensus will be achieved
%without using slower synchronous calls. For the tight timing we wish to
%support, we need guarantees of when a consistent state has been achieved.
%Second, these systems also all assume that updates are initiated from a single
%client process rather than a parallel set of processes as is the standard in
%HPC environments.
%
%\DDT uses a second layer of coordination on the client side that greatly
%increases the scalability by consolidating messages from clients into unique
%sets prior to sending to the overall coordinator. A gossip
%protocol~\cite{ganesh:2003:gossip-protocols} may appear sufficient for this
%purpose, but the delay of eventual consistency is strictly avoided with this
%protocol to ensure guarantees at particular states in the code. For example, if
%a failure occurs, the global understanding of the role of all processes is
%required in order for effective communication to occur for operations like
%creating sub-transactions or voting. In this case, the protocol can offer
%stronger statements about consistency than these protocols offer.  These
%features offer a way to easily scale the transaction protocol given the
%guarantees we wish to offer. The IOD approach does nothing to address these
%concerns.
%
%Another effort to offer consistency and data integrity for the ZFS file
%system~\cite{zhang:2010:zfs} covers some of the same territory. Instead of a
%focus on the processes all having a notion of completion as a transaction, this
%work focuses on the integrity of the data movement operations. We view this
%work as something that should be considered hand-in-hand with a transaction
%approach to ensure the integrity of the movement of the data in addition to the
%agreement of processes about the successful completion of a parallel operation.

@@

\subsection{ACG}\label{acg}

>>>>

The complete design seeks to offer high availability, byte-granular,
multi-version concurrency control. Through the use of a copy-on-write style
mechanism reducing space and time requirements, multiple versions of an object
can be stored efficiently. By assuming the client interface will be through an
IO library, a more complicated interface that offers richer functionality can
be incorporated while requiring only minimal end-user code changes.  Managing
most data access in a platform-local layer rather than requiring writing to
centralized storage will better support the performance and energy requirements
of extreme scale integrated application workflows.

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=\columnwidth]{images/arch-mapping}
\caption{Target Architecture and Component Mapping}
\label{fig:arch-mapping}
\end{figure}

At a more detailed view, the various layers of the IO stack each contribute
different functionality and performance implications.  The
architecture (Figure~\ref{fig:arch-mapping}) incorporates five layers, some of
which have potentially optional components.  The top layer is generally a high
level IO library, such as the demonstration HDF-5 library~\cite{hdf5}. It is in
dark blue. The intent is to only have access to the storage stack through such
an API to manage the complexity of working with the lower layers and to enable
advanced functionality that would require more direct user intervention. This
layer incorporate the necessary features for ACGs from a end-user's
perspective.

Below the user API is an IO forwarding layer that redirects IO calls from the
compute nodes to the IO dispatching layer (in black).  This IO forwarding layer
is analogous to the function of the IO nodes in a BlueGene machine or the
passive data staging processes demonstrated
previously~\cite{nisar:2008:staging,Abbasi:2009:datatap}. The next two layers
have considerable functionality. The IO dispatcher (IOD) serves as the primary
storage interface for the IO stack and is the only way to reach the persistent
storage array in lower layers (in green). Ideally, the IOD layer's
functionality can be optional based on available hardware. Much of the
functionality offered at this layer would shift either up or down the stack as
discussed in detail below. The Distributed Asynchronous Object Storage (DAOS)
layer serves as the persistent storage interface and is intended to be the
traditional file system-like foundation on which everything else is built with
no dependence on any technologies specified above it (in pink and yellow). For
example, the IOD layer with or without burst buffers is not required for DAOS
to operate properly.  Instead, the DAOS layer can handle all of the IO
operations from the user API layer, albiet with the potential performance
penalty of manipulating the shared, persistent storage array. At the bottom is
the Versioning Object Storage Device (VOSD) (in purple).  It serves as the
interface for storing objects of all types efficiently. Think of this layer as
the physical disk interface layer. In terms of Lustre, this would replace the
file system on individual storage targets with an interface friendlier to the
containers of objects concept used in the higher layers.

Along with the analysis of the published design documents, a discussion of the
design philosophy representing the overall intent is presented. This
information represents information that may or may not have been written down,
but is the intent of ultimate product.  These ideas are presented to give a
fuller picture of where the project is going rather than dwelling on any
limitations of the published documents.

@@

The major previous work is really limited to full file systems rather than the
mountain of file system adjustments made over the years.

Ceph~\cite{weil:ceph} is a distributed object store and file system. It offers
both a POSIX and object interface including features typically found in parallel
file systems. Ceph's unique striping approach uses pseudo-random numbers with a
known seed eliminating the need for the metadata service to track where each
stripe in a parallel file is placed.

PVFS~\cite{carns:pvfs} offers optimizations to reduce metadata server load,
such as a single process opening a file and sharing the handle.

Lustre~\cite{braam:lustre-arch} has become the defacto standard on most major
clusters offering scalable performance and fine-grained end-user and programmatic control over how data is placed in the storage system. Scalability limitations are part of the motivation for the FFSIO project.

GPFS~\cite{schmuck:gpfs} offers a hands-off approach for providing good performance for scaling parallel IO tasks and is used extensively by its owner, IBM.

Panasas~\cite{panasas:architecture} seeks to offer a dynamically adaptive striping system that detects the need for additional stripes for performance and adjusts the file layout as necessary.

Other file systems, like GoogleFS~\cite{ghemawat:googlefs}, address distributed
rather than parallel computing and cannot be compared directly. The primary
difference between distributed and parallel file systems is the ability of the
file system to retrieve data simultaneously from multiple clients, in parallel,
and treat the resulting collection of pieces as a single object. Distributed
file systems rely on a single client creating a file, but distributing the set
of files across a wide array of storage devices. The other, popular distributed
file system of note is HDFS~\cite{Shvachko:2010:hdfs} that is distributed as
part of Hadoop. These other file systems are mainly of interest in the context
of the ACG features of FFSIO and will be discussed more in
Section~\ref{sec:acg}.

@@

\item
  \textbf{File}: In the HDF5 data model the container of an HDF5 infoset
  is represented by a file. It is a collection of objects that also
  explains the relationship between them. Every file begins with a root
  group ``/'', which serves as the ``starting-point'' in the object
  hierarchy.

@@

\section{IO Dispatcher Layer}
\label{sec:iod}

@@

@@

The second iteration of the protocol~\cite{lofstead:2013:pdsw-txn} fixed
scalability issues and demonstrated a scalable client-side coordination model
with excellent performance. The performance measured for a complex transaction
with \DDT is illustrated in Figure~\ref{fig:performance}. This performance is
explored in detail in a previous paper~\cite{lofstead:2013:pdsw-txn}.

@@

@@

@@

\subsection{ACG}
\label{sec:acg}
